### 2007-01-15 Documentation/appendix.tex
*** empty log message ***
JK - Minor modifications in the appendices (except appendix B)

Besides the usual things like a working libc you need to have
 \begin{itemize}
        \item[gcc]      The GNU compiler collection\\
                        \verb+gcc.gnu.org+
        \item[GSL]      The GNU Scientific Library\\
                        \verb+www.gnu.org/software/gsl/+
\end{itemize}
The library \verb+libglobes+ should in principle compile with any C/C++
compiler but the \verb+globes+ binary uses the \verb+argp+ facility of \verb+glibc+
to parse its command line options. However, on platforms where \verb+argp+
is lacking \GLOBES\ has replacement code, thus it should also work
there. \GLOBES\ is, however, using the C99 standard in order to handle
complex numbers, but that is the only feature of C99 used.

GSL is also available as rpm+s from the various distributors of
GNU/Linux, see their web sites for downloads. Chances are that gcc and
GSL are already part of your installation.  For building \GLOBES\ from
source, however, not only working libraries for above packages are
needed but also the headers, especially for GSL. Depending on your
installation, eg. on RedHat/Fedora, of GSL this may require to
additionally install a rpm-package named \verb+gsl-devel+. If GSL has been
installed from the tar-ball as provided by gnu.org no problems should
occur. Furthermore you need a working \verb+make+ to build and install
\GLOBES.

@@

\GLOBES\ follows the standard GNU installation procedure.  To compile
\GLOBES\ you will need an ANSI C-compiler.  After unpacking the
distribution the Makefiles can be prepared using the configure
command,\\
\verb+  ./configure+\\
You can then build the library by typing,\\
\verb+  make+\\
A shared  version of the library will be compiled by
default.

@@

Under certain circumstances it may be useful to use a static version of
libglobes or any of the binaries, e.g. when running on a cluster.

The \verb+configure+ script accepts the option \verb+--disable-shared+, in
which case only static objects are built, i.e. only a static
version of libglobes. In case your system does not support shared
libraries the \verb+configure+ script recognizes this. If you give no
options to \verb+configure+, both shared and static versions are built
and will be installed. All binaries, however, will use dynamic
linking. If you want to build static binaries, use
LDFLAGS=+-all-static+ for building them.

Sometimes it is convenient, eg. for debugging purposes, to have a
statically linked version of a program using \GLOBES\, which is easiest
achieved by just linking with \verb+libglobes.a+. If you need a completely
statically linked version, please, have a look at the Makefile in the
\verb+examples+ directory.\\
\verb+  make example-static+\\ 
produces a statically linked program that should in principle run on
most Linuxes.  It should be straightforward to adapt this example to
your needs.

All these options rely on a working \verb+gcc+ installation. It seems that
\verb+gcc 3.x+ is broken in subtle way which makes it necessary to add a
symbolic link in the gcc library directory. The diagnostic for this
requirement is that building static programs fails with the error
message \verb+cannot find -lgcc_s+. In those cases, find \verb+libgcc.a+ and add
a symbolic link in the same directory where you found it (this
requires probably root privileges)\\
\verb+  ln -s libgcc.a libgcc_s.a+

@@

In order to fully exploit the functionality offered by Condor one
should submit the jobs into the so called 'standard universe'. To do
this, it is necessary to re-link the application with the
Condor-library (this assumes that Condor is installed)\\ 
\verb+  condor_compile gcc your_object_files -static `globes-config --libs`+\\
It may be necessary to prefix the call of \verb+globes-config+ with
the path to it, in case that location is not in \verb+$PATH+.
\index{norm}{Condor|)}

@@

Sometimes the GNU scientific library is not available or is installed
in a non-standard location. This situation can arise in an
installation without root privileges. In this case one can specify
\verb+--with-gsl-prefix=path_to_gsl+ as option to the \verb+configure+ script.
If one wants to use a shared version of \verb+libgsl+ then one has to make
sure that the linker find the library at run-time. This can be
achieved by setting the environment variable \verb+LD_LIBRARY_PATH+
correctly, i.e. (in bash)\\
\verb+  export LD_LIBRARY_PATH='path_to_gsl'+\\
You also can use a static version of GSL by either building \GLOBES\
with \verb+LDFLAG='-all-static'+ or by configuring GSL with
\verb+--disable-shared+. In both cases no further actions like setting any
environment variables is necessary.

@@

The standard rpm-based installation of GSL does not provide any header
files for GSL, which are however needed to compile \GLOBES\. You have to
install an additional rpm-package called \verb+gsl-devel+. Alternatively
you can install GSL from a tar-ball and use the \verb+--with-gsl-prefix+
option to the configure script of \GLOBES.

@@

   By default, \verb+make install+ will install the package+s files in
\verb+/usr/local/bin+, \verb+/usr/local/man+, etc.  You can specify an
installation prefix other than \verb+/usr/local+ by giving \verb+configure+ the
option \verb+--prefix=PATH+.

   You can specify separate installation prefixes for
architecture-specific files and architecture-independent files.  If you
give \verb+configure+ the option \verb+--exec-prefix=PATH+, the package will use
PATH as the prefix for installing programs and libraries.
Documentation and other data files will still use the regular prefix.

@@

   Some packages pay attention to \verb+--enable-FEATURE+ options to
\verb+configure+, where FEATURE indicates an optional part of the package.
They may also pay attention to \verb+--with-PACKAGE+ options, where PACKAGE
is something like \verb+gnu-as+ or \verb+x+ (for the X Window System).  The
\verb+README+ should mention any \verb+--enable-+ and \verb+--with-+ options that the
package recognizes.

@@

This feature is experimental and your mileage my vary!

@@

The trick here, is that we use SWIG (\verb+www.swig.org+) to generate
a wrapper file for \GLOBES. The wrapper file is part of the \GLOBES\
tar-ball (\verb+globes/globes_perl.c+) and hence you should not need SWIG to
be installed on your system.

@@

This feature is experimental and your mileage my vary!

Many people find binary RPMs useful, therefore we provide an optional
feature \verb+--enable-rpm-rules+ which should produce all the necessary
makefile rules for RPM building. To actually build RPMs requires that
your system is properly setup for that. How you can do that you can
learn at \verb+http://www.rpm.org+. You then can use \verb+make rpm+, most likely
you will need to be root to do that (\verb+sudo+ won't work!).

@@

   There may be some features \verb+configure+ can not figure out
automatically, but needs to determine by the type of host the package
will run on.  Usually \verb+configure+ can figure that out, but if it prints
a message saying it can not guess the host type, give it the
\verb+--host=TYPE+ option.  TYPE can either be a short name for the system
type, such as \verb+sun4+, or a canonical name with three fields\\
\verb+  CPU-COMPANY-SYSTEM+

@@

A standard neutrino factory scenario can be simulated with the file {\tt NFstandard.glb}. The basic version was
used within \cite{Huber:2002mx} (NuFact-II scenario). This reference should be cited if the file {\tt
NFstandard.glb} is used for a scientific publication or a talk. For calculations that involve {\tt
NFstandard.glb}, the following additional files are required:
\bi
\item {\tt XCC.dat} (charged current cross sections)
\item {\tt XNC.dat} (neutral current cross sections)
\ei
The neutrino beam is produced by the decay of muons stored in a storage ring at a parent energy of
$\mathrm{E_\mu = 50\,GeV}$. A number of $1.06\cdot10^{21}$ useful muon decays per year are assumed in each
polarity (corresponding to $5.3\cdot10^{20}$ useful muon decays per year and polarity for simultaneous operation
with both polarities) and 4~years $\nu$-running and 4~years $\bar{\nu}$-running is
assumed. The fiducial mass of the MID detector is taken to be $\mathrm{m_{det} = 50 \,kt}$ at a baseline of
L~=~3000~km. The energy resolution is $\mathrm{\sigma_e=15\%\cdot E}$. The following rules are defined within
{\tt NFstandard.glb}:
\begin{center}
\begin{tabular}{|l|ll|c|c|}
\hline \hline
\multicolumn{3}{|l|}{Disappearance -- $\mu^+$-stored} & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$ \\ \hline
Signal & $0.35 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$ & & 0.001 & $10^{-4}$ \\
 & & & & \\
Background & $1.0\cdot 10^{-5} \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & & 0.2 & $10^{-4}$ \\ \hline \hline
\multicolumn{3}{|l|}{Appearance -- $\mu^+$-stored} & & \\ \hline
Signal &  $0.45 \, \otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{CC}$ & & 0.001 & $10^{-4}$ \\
 & & & & \\
Background &  $5.0\cdot 10^{-6} \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_x)_\mathrm{NC}$ &  $5.0\cdot
10^{-6} \, \otimes \, (\bar{\nu}_\mu \rightarrow\bar{\nu}_\mu)_\mathrm{CC}$ & 0.2& $10^{-4}$\\ \hline \hline
\multicolumn{3}{|l|}{Disappearance -- $\mu^-$-stored} & & \\ \hline
Signal &  $0.45 \, \otimes \, (\nu_\mu \rightarrow \nu_\mu)_\mathrm{CC}$ & & 0.001& $10^{-4}$\\
 & & & & \\
Background &  $1.0\cdot 10^{-5} \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & & 0.2& $10^{-4}$\\ \hline \hline
\multicolumn{3}{|l|}{Appearance -- $\mu^-$-stored} & & \\ \hline
Signal & $0.35 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$  & & 0.001& $10^{-4}$\\
 & & & & \\
Background &  $5.0\cdot 10^{-6} \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & $5.0\cdot 10^{-6} \, \otimes \, (\nu_\mu \rightarrow
\nu_\mu)_\mathrm{CC}$  & 0.2& $10^{-4}$\\ \hline \hline
\end{tabular}
\end{center}

@@

A common issue with \GLOBES\ is confusion about the proper units for
the input flux files for use in \AEDL\ experiment descriptions. Source of
the confusion is an undocumented factor 5.2 in \GLOBES\ versions older
than 3.0 the fluxes are multiplied with (see below). In Version 3.0 and 
higher, the alternative flux environment {\tt nuflux} is provided, 
which does not contain this factor. The following material is
based on the old environment {\tt flux}. For the use of {\tt nuflux},
replace the factor 5.2 by unity.

@@

This replacement is not necessary for neutrino factory built-in fluxes,
and built-in beta beams fluxes where not supported by earlier versions
of \GLOBES .

>>>>

Besides the usual things like a working libc you need to have
 \begin{itemize}
        \item[gcc]      The GNU compiler collection\\
                        \verb+gcc.gnu.org+
        \item[GSL]      The GNU Scientific Library\\
                        \verb+www.gnu.org/software/gsl/+
\end{itemize}
The library \verb+libglobes+ should in principle compile with any C/C++
compiler but the \verb+globes+ binary uses the \verb+argp+ facility of \verb+glibc+
to parse its command line options. However, on platforms where \verb+argp+
is lacking, \GLOBES\ has replacement code, thus it should also work
there. \GLOBES\ is, however, using the C99 standard in order to handle
complex numbers, but that is the only feature of C99 used.

GSL is also available as rpm's from the various distributors of
GNU/Linux, see their web sites for downloads. Chances are that gcc and
GSL are already part of your installation.  For building \GLOBES\ from
source, however, not only working libraries for the above packages are
needed, but also the headers, especially for GSL. For some installations
of GSL, eg. on RedHat/Fedora, this may require to
additionally install a rpm-package named \verb+gsl-devel+. If GSL has been
installed from the tar-ball as provided by gnu.org, no problems should
occur. Furthermore you need a working \verb+make+ to build and install
\GLOBES.

@@

\GLOBES\ follows the standard GNU installation procedure.  To compile
\GLOBES\ you will need an ANSI C-compiler.  After unpacking the
distribution, the Makefiles can be prepared using the configure
command,\\
\verb+  ./configure+\\
You can then build the library by typing,\\
\verb+  make+\\
A shared  version of the library will be compiled by
default.

@@

Under certain circumstances it may be useful to use a static version of
libglobes or any of the binaries, e.g.\ when running on a cluster.

The \verb+configure+ script accepts the option \verb+--disable-shared+, in
which case only static objects are built, i.e.\ only a static
version of libglobes. In case your system does not support shared
libraries the \verb+configure+ script recognizes this. If you give no
options to \verb+configure+, both shared and static versions are built
and will be installed. All binaries, however, will use dynamic
linking. If you want to build static binaries, use
\verb+LDFLAGS='-all-static'+ for building them.

Sometimes it is convenient, eg. for debugging purposes, to have a
statically linked version of a program using \GLOBES, which is easiest
achieved by just linking with \verb+libglobes.a+. If you need a completely
statically linked version, please, have a look at the Makefile in the
\verb+examples+ directory.\\
\verb+  make example-static+\\ 
produces a statically linked program that should in principle run on
most Linuxes.  It should be straightforward to adapt this example to
your needs.

All these options rely on a working \verb+gcc+ installation. It seems that
\verb+gcc 3.x+ is broken in a subtle way which makes it necessary to add a
symbolic link in the gcc library directory. The diagnostic for this
requirement is that building static programs fails with the error
message \verb+cannot find -lgcc_s+. In those cases, find \verb+libgcc.a+ and add
a symbolic link in the same directory where you found it (this
requires probably root privileges)\\
\verb+  ln -s libgcc.a libgcc_s.a+

@@

In order to fully exploit the functionality offered by Condor one
should submit the jobs into the so called 'standard universe'. To do
this, it is necessary to re-link the application with the
Condor-library (this assumes that Condor is installed)\\ 
\verb+  condor_compile gcc your_object_files -static `globes-config --libs`+\\
It may be necessary to prefix the call of \verb+globes-config+ with
the path to it, in case that this location is not in \verb+$PATH+.
\index{norm}{Condor|)}

@@

Sometimes, the GNU scientific library is not available or is installed
in a non-standard location. This situation can arise in an
installation without root privileges. In this case, one can specify
\verb+--with-gsl-prefix=path_to_gsl+ as option to the \verb+configure+ script.
If one wants to use a shared version of \verb+libgsl+ then one has to make
sure that the linker can find the library at run-time. This can be
achieved by setting the environment variable \verb+LD_LIBRARY_PATH+
correctly, i.e. (in bash)\\
\verb+  export LD_LIBRARY_PATH='path_to_gsl'+\\
You also can use a static version of GSL by either building \GLOBES\
with \verb+LDFLAG='-all-static'+ or by configuring GSL with
\verb+--disable-shared+. In both cases no further actions like setting any
environment variables is necessary.

@@

The standard rpm-based installation of GSL does not provide any header
files for GSL, which are however needed to compile \GLOBES. You have to
install an additional rpm-package called \verb+gsl-devel+. Alternatively
you can install GSL from a tar-ball and use the \verb+--with-gsl-prefix+
option to the configure script of \GLOBES.

@@

   By default, \verb+make install+ will install the package's files in
\verb+/usr/local/bin+, \verb+/usr/local/man+, etc.  You can specify an
installation prefix other than \verb+/usr/local+ by giving \verb+configure+ the
option \verb+--prefix=PATH+.

   You can specify separate installation prefixes for
architecture-specific files and architecture-independent files.  If you
give \verb+configure+ the option \verb+--exec-prefix=PATH+, the package will use
\verb+PATH+ as the prefix for installing programs and libraries.
Documentation and other data files will still use the regular prefix.

@@

   Some packages pay attention to \verb+--enable-FEATURE+ options to
\verb+configure+, where \verb+FEATURE+ indicates an optional part of the package.
They may also pay attention to \verb+--with-PACKAGE+ options, where \verb+PACKAGE+
is something like \verb+gnu-as+ or \verb+x+ (for the X Window System).  The
\verb+README+ should mention any \verb+--enable-+ and \verb+--with-+ options that the
package recognizes.

@@

This feature is experimental and your mileage may vary!

@@

The trick here is, that we use SWIG (\verb+www.swig.org+) to generate
a wrapper file for \GLOBES. The wrapper file is part of the \GLOBES\
tar-ball (\verb+globes/globes_perl.c+) and hence you should not need SWIG to
be installed on your system.

@@

This feature is experimental and your mileage may vary!

Many people find binary RPMs useful, therefore we provide an optional
feature \verb+--enable-rpm-rules+ which should produce all the necessary
Makefile rules for RPM building. To actually build RPMs requires that
your system is properly setup for that. You can learn how to do that
at \verb+http://www.rpm.org+. You then can use \verb+make rpm+, most likely
you will need to be root to do that (\verb+sudo+ won't work!).

@@

   There may be some features \verb+configure+ can not figure out
automatically, but needs to determine by the type of host the package
will run on.  Usually \verb+configure+ can figure that out, but if it prints
a message saying it can not guess the host type, give it the
\verb+--host=TYPE+ option. \verb+TYPE+ can either be a short name for the system
type, such as \verb+sun4+, or a canonical name with three fields\\
\verb+  CPU-COMPANY-SYSTEM+

@@

A standard neutrino factory scenario can be simulated with the file {\tt NFstandard.glb}. The basic version was
used within \cite{Huber:2002mx} (NuFact-II scenario). This reference should be cited if the file {\tt
NFstandard.glb} is used for a scientific publication or a talk. For calculations that involve {\tt
NFstandard.glb}, the following additional files are required:
\bi
\item {\tt XCC.dat} (charged current cross sections)
\item {\tt XNC.dat} (neutral current cross sections)
\ei
The neutrino beam is produced by the decay of muons stored in a storage ring at a parent energy of
$\mathrm{E_\mu = 50\,GeV}$. A number of $1.06\cdot10^{21}$ useful muon decays per year are assumed in each
polarity (corresponding to $5.3\cdot10^{20}$ useful muon decays per year and polarity for simultaneous operation
with both polarities) and 4~years $\nu$-running and 4~years $\bar{\nu}$-running is
assumed. The fiducial mass of the MID detector is taken to be $\mathrm{m_{det} = 50 \,kt}$ at a baseline of
L~=~3000~km. The energy resolution is $\mathrm{\sigma_e=15\%\cdot E}$. The following rules are defined within
{\tt NFstandard.glb}:
\begin{center}
\begin{tabular}{|l|ll|c|c|}
\hline \hline
\multicolumn{3}{|l|}{Disappearance -- $\mu^+$-stored} & $\sigma_\mathrm{norm}$ & $\sigma_\mathrm{cal}$ \\ \hline
Signal & $0.35 \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$ & & 0.025 & $10^{-4}$ \\
 & & & & \\
Background & $1.0\cdot 10^{-5} \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_x)_\mathrm{NC}$ & & 0.2 & $10^{-4}$ \\ \hline \hline
\multicolumn{3}{|l|}{Appearance -- $\mu^+$-stored} & & \\ \hline
Signal &  $0.45 \, \otimes \, (\nu_e \rightarrow \nu_\mu)_\mathrm{CC}$ & & 0.025 & $10^{-4}$ \\
 & & & & \\
Background &  $5.0\cdot 10^{-6} \, \otimes \, (\bar{\nu}_\mu \rightarrow \bar{\nu}_x)_\mathrm{NC}$ &  $5.0\cdot
10^{-6} \, \otimes \, (\bar{\nu}_\mu \rightarrow\bar{\nu}_\mu)_\mathrm{CC}$ & 0.2& $10^{-4}$\\ \hline \hline
\multicolumn{3}{|l|}{Disappearance -- $\mu^-$-stored} & & \\ \hline
Signal &  $0.45 \, \otimes \, (\nu_\mu \rightarrow \nu_\mu)_\mathrm{CC}$ & & 0.025& $10^{-4}$\\
 & & & & \\
Background &  $1.0\cdot 10^{-5} \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & & 0.2& $10^{-4}$\\ \hline \hline
\multicolumn{3}{|l|}{Appearance -- $\mu^-$-stored} & & \\ \hline
Signal & $0.35 \, \otimes \, (\bar{\nu}_e \rightarrow \bar{\nu}_\mu)_\mathrm{CC}$  & & 0.025& $10^{-4}$\\
 & & & & \\
Background &  $5.0\cdot 10^{-6} \, \otimes \, (\nu_\mu \rightarrow \nu_x)_\mathrm{NC}$ & $5.0\cdot 10^{-6} \, \otimes \, (\nu_\mu \rightarrow
\nu_\mu)_\mathrm{CC}$  & 0.2& $10^{-4}$\\ \hline \hline
\end{tabular}
\end{center}

@@

A common issue with \GLOBES\ is confusion about the proper units for
the input flux files for use in \AEDL\ experiment descriptions. Source of
the confusion is an undocumented factor 5.2 with which the fluxes
are multiplied in \GLOBES\ versions older than 3.0 (see below). In Version 3.0 and 
higher, the alternative flux environment {\tt nuflux} is provided, 
which does not contain this factor. The following material is
based on the old environment {\tt flux}. For the use of {\tt nuflux},
replace the factor 5.2 by unity.

@@

This replacement is not necessary for neutrino factory built-in fluxes,
and built-in beta beam fluxes were not supported by earlier versions
of \GLOBES .
